### 2019-07-10 manuscript/response.tex

\renewenvironment{quote}{\bigskip\noindent\itshape\ignorespaces}{\smallskip}

\begin{document}

\hfill \today

Dear Dr. Schoen and reviewers,

\medskip
Thank you very much for your close reading of our manuscript and supportive comments. We appreciate your constructive suggestions, each of which we address below. You will find a big difference on the number of  models  that we have fitted for the revision based on your suggestions (originally 12 now 26). The two main reasons to increase the number of models were to first address one key question by Referees \#2 and \# 3 about the symmetry of the hidden states, and second to address a key question about more complex model. For the symmetry of the hidden states we found that is relevant to have asymmetric hidden state rates and in our case, it strengthened the differences amongst net diversification rates. The second question,  comes from a vague idea that we had while submitting the first version of this article: by increasing from two to three states the models, are we explaining better the history of diversification? This second question could not be addressed directly by model comparison statistics because the two-state and three-state models receive different input (i.e. (0,1) versus (0,1,2)), thus different data. However, the species and the tree are the same and should be delivering the same evidence despite any particular coding. During this revision, we figured a way to recode the three-state models to make them equivalent to the two-state models (the lumped models). Thanks to this equivalency, two-state and three-state models can now be compared to argue for the importance of adding key trait information and states in diversification analyses.

We hope the changes and the new models are satisfactory and exciting to you and the reviewers.

@@

We reworded this whole paragraph, as part of better explaining how polyploidization disables gametophytic SI.

\begin{quote}
--83 Nitpicky, but is it accurate to refer to this as a false-positive problem?
``False positive'' implies to me the erroneous rejection of the null, but in this case the situation is more of the null being mis-specified with respect to the question the researcher is after.
The null---no rate variation---is correctly rejected.
This section probably should include a reference to Rabosky and Goldberg.
\end{quote}

We reworded this whole paragraph, as part of better explaining the SSE family of models.

\begin{quote}
--115 I'm uncomfortable with this way of handling multiple cytotypes.
The most commonly reported cytotype could easily not be the most common cytotype in the wild (I assume that most taxa are represented by relatively few counts), and regardless, there is biology here.
Specifically, this methodology would artificially reduce the number of ploidy changes (if a taxon is polymorphic for ploidy, there's at least one event there), and the omitted changes would be disproportionately those towards the tips of the phylogeny, so not only would rates be depressed, but the bias is of a sort that the estimation of diversification parameters is particularly sensitive to (the authors will understand why this is so).
\end{quote}
We acknowledge this is potentially a problem. It is true that chromosome number multiplicity can bias towards more (or less) transitions because it heavily depends on the most common cytotype. For the majority of species (~85\%)  the ploidy was not directly dictated by us but by the original sources of the data. In the originally sources ploidy has been classified not only most common cytotype but geographical distribution, multivalence, and/or  genome sizes. The remaining 15\% is what we assigned based on genus or family cytotype multiplicity. We cross-validated all of the assignments with the data from \citet{robertson_2011} where ploidy was assigned via a chromosome number change model using ChromEvol \citep{mayrose_2010}. We differ in exactly 35 species from \citet{robertson_2011} were we found a new reference or a new chromosome count. Classifying using a chromosome number change model only could also suffer from inflation of polyploidy events due to the lack of diploidization parameters.

@@

Fixed.

\begin{quote}
--140 I lost track a little of how the data were coded -- maybe add an explicit statement as to how many tips had both forms of data, how many had only breeding system info, how many had only ploidy info (I see now, after the fact, that those data are included in the figure), and how missing data were handled
\end{quote}

Fixed.

\begin{quote}
--It's not intuitive to me what the effects of forcing the hidden state change rates to be symmetrical are.
Did the authors investigate this?
Or has it been explored in the literature?
Similarly, it seems to me that fixing the transition rates of the observed characters to be independent of the hidden state could do weird things---the point of these SSE models is that state change rates and state-dependent diversification rates need to be co-estimated
\end{quote}
We appreciate this suggestion and we followed. The main result after fitting models with symmetrical vs. asymmetrical rates is that the direction of the results does not change but the amount of difference amongst net diversification rates does. We found using bayes factors that a model including asymmetrical hidden rates is preferred over a model with equal hidden rates. Also,  the asymmetry in the trait rates did not add much to the differences according to the bayes factors. These results can be found on table and figures %TODO put number.

@@

\begin{quote}
--221  Propagation
\end{quote}

Fixed.

\begin{quote}
--I would be more comfortable if each model was run at least twice, independently, to better assess whether the MCMC has indeed converged
\end{quote}

Ouch.  Rosana? % TODO: Still need to calculate some stats here

\begin{quote}
--the use of run time as a description of the analysis seems odd to me.
As a reader, I'm not interested in the power of a given computer -- I want to know how many generations the MCMC was run for, what the ESSs are, etc.
\end{quote}
We ran each model for 200,000 generations. For the main models with asymmetrical hidden rates all of the parameters have ESS of 300. For the most difficult models were all rates were different  the minimum is 125. Unfortunately, we don't have access to more of 96 hours of computational time to increase those but we have calculated the convergence as indicated above.

@@

\begin{quote}
--beginning of the Discussion is perhaps unnecessary -- it's more like Introduction material
\end{quote}

consider % TODO

\begin{quote}
--are the results from the analyses that include diploidization coherent?
I'm not sure how to phrase this -- I guess I'm asking if there's any suggestion that that parameterization is indeed capturing the biological phenomenon of diploidization, or is it just soaking up noise?
\end{quote}

The soaking noise comes from the hidden state so in the presence of hidden states finding that diploidization matters is key. All the bayes factors indicate strong support for models that include diploidization. However, the diploidzation parameter per se is difficult to estimate from ploidy data because of the quality of the classification. For those two reasons we are conservative about overstating the importance of diploidization.

@@

We're pleased you found the results interesting!

\begin{quote}
I have two ``major'' comments.
The first is relatively easy to fix and is linked to the writing/structure of the paper.
I often had the impression that too much knowledge is assumed from the reader throughout the manuscript.
For instance, more information should be given in the intro regarding why self-incompatibility breakdowns with polyploidy (it is briefly explained, but only later in the manuscript).
Moreover, some methodological approaches could be better explained.
In particular, HISSE models are relatively recent and many readers may not know them well.
I thus think that it would be important to describe them better and also better explain the results (using biological terms) obtained.
I mention some examples below in the detailed notes where I think explanations could be added.
\end{quote}

We have rewritten parts of the manuscript, especially in the Introduction, to try to explain better the relationship between polyploidy and self-incompatibility, and the interpretations of hidden state models.
% E: I labeled the two paragraphs that I modified heavily.

\begin{quote}
My second comment has to do with the parameterization of the models.
Overall, I think that the authors have done a great job in justifying their models and explaining them.
I think it is important to mention.
In addition, their implementation of the models in revBayes is noteworthy and although not stated precisely in the manuscript, I believe this represents a great scientific contribution. 
Nevertheless, I question the decision to restrict the transition rates to and from the hidden states to be all equal.
Indeed, we have no idea that could make us think that this would be the case.
The authors justify this by ``Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.''
I am not completely convinced by this argument.
First, constraining transition rates can have an impact on the estimation of diversification rates.
Second, specifically because the authors are not interested in estimating the rates themselves, they should not be too worried about increasing the number of parameters.
This is especially true in a Bayesian framework where it is sometimes better to have overparameterized models than under-parameterized models (see Lemmon and Moriarty. 2004. Systematic Biology 53:265–277).
I think it would have been important to at least present the results of an unconstrained model to show that this decision to constrain the rates do not affect the conclusions.
\end{quote}
We have now fitted models with asymmetrical hidden rates and asymmetrical trait rates. We found that the former are important and strengthened the differences. We found no evidence that having all asymmetrical rates is preferable over just asymmetrical hidden for our specific example. (Figs and table)

@@

I have a few comments that I hope the authors will find useful.
\end{quote}

Thank you for this positive summary.

\begin{quote}
The first relates to the use of the hidden states in the models.
I agree completely with the rationale and meaning of including hidden states.
However, through my reading of the manuscript, I interpret the authors as using them simply to account for variation within a particular character state combination.
In other words, the authors are working under an assumption of trait-dependent diversification, with the hidden states simply providing a more nuanced view of the process.
But the authors are not considering a completely alternative view, which is that none of these trait combinations are strong drivers of diversification and that the shifts being inferred are simply coincidental.
Such an interpretation could be ruled out by explicitly fitting a model of trait-independent diversification.
This is achieved by linking the diversification rate parameters among the different character state combinations -- i.e., $ID_A=CD_A=CP_A$, and $ID_B=CD_B=CP_B$.
In other words, because the diversification rate differences are completely linked among the possible states, but unlinked by the hidden states, this forces an explicitly trait-independent model.
The fit of such a model could be included in the set of models in Table 1 and compared using Bayes Factors.
Of course, admittedly, I am not familiar with revBayes and how the model space is set up, so it's possible this issue is already being addressed.
But, if so, some clarity in the Methods/Results is warranted in how trait-independent models were evaluated (or even ruled out) in the analyses presented.
\end{quote}

We have fitted the trait independent models (character-independent CID). We appreciated this suggestion because  we hadn't considered it and it allowed us to parse out the effect of the traits in the diversification process. By comparing bayes factors, models with traits are strongly preferred over CID (table )
%TODO: write table number

@@

\begin{quote}
Line 465-469: This sentence was a bit difficult to grasp during my first reading of it.
I get the point, but I think the authors might consider revising it for clarity.
\end{quote}

Rephrased. % TODO

\begin{quote}
Line 476-480: The type I error issues reported by Rabosky and Goldberg (2015) are, in my view, fairly universal.
So, I think it is safe to say that MuSSE ``does suffer'' issues of model misspecification/type I error.
Also, it is worth noting that there are several implementations that already include hidden states in a MuSSE model, most notably the HORRIBLY named SecSSE model, as well as the models used in a recent empirical paper on diatoms by Nakov et al (which I've seen on bioRxiv, though might be published by now).
\end{quote}

Yes, and indeed this has already been shown.
We fixed the text and added a reference. Nakov paper has not been accepted yet for publication or it has not been announced (July 9th, 2019).

### 2019-07-10 manuscript/methods.tex
Editing methods and adding the SSE models fig

Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011} and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus or based on SI/SC classification.
For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid. 
Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.

@@

Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 595 species with ploidy and/or breeding system information on the tree.
% FIXME: Rosana will update numbers around here
Of these, 290 had information for both states.
The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
We retained all 595 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
A total of 405 taxa without any information about breeding system or ploidy were excluded.
%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
% R-This is very important but I have removed it because I think it is a full on paper by its own
%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.

@@

% E: Rosana, please check this paragraph.  Did I get it correct?
To test if diversification rate heterogeneity is due only to a hidden trait and not the ploidy trait, we used character-independent diversification models \citep[CID,][]{beaulieu_2016}.
Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with polyploidization ($\rho$) and transitions between hidden states ($\alpha$).

@@

To parse out the effect of breeding system on diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of the breeding system states into hidden states $A$ and $B$.
This $I/C+A/B$ breeding system and hidden state model has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
We also considered more complex models with different transition rates in each hidden state (???) and asymmetric transitions between hidden states ($\alpha$ and $\beta$). % FIXME
Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.

@@

Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association (see the Introduction).
Therefore, we considered a multi-state model that investigates the contribution of both traits and the allowable transitions between them.
This $ID/CD/CP$ ploidy and breeding system model has states for self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$; self-incompatible polyploids do not occur, as explained in the Introduction.
There are six parameters for state-specific diversification rates ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation; $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.

% E: This paragraph is new.  Rosana, please check.
We would like to test whether including an additional trait significantly improves the model fit.
For example, does the model with both breeding system and ploidy provide a significantly better fit than the model with only ploidy?
To answer this, it is not valid to compare the $IC/CD/CP$ and $D/P$ model fits, as defined above, because they use different data.
Similarly, the $IC/CD/CP$ and $I/C$ models cannot be compared statistically.
Instead, we define lumped models \citep{tarasov_2019} that use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models. % should ref a figure

% E: We now have so many models to describe, I'm feeling like all this text may not be the best way.  I'm wondering if we can rely on figures/tables instead, and only write out the overall strategy.

Although the $ID/CD/CP$ model allows for two traits to potentially influence diversification, other unknown factors could as well.
We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, huang_2018}.
A fully parameterized version of this $ID/CD/CP+A/B$ model would have 26 rate parameters. 
Because our goal was to look for diversification rate differences associated with ploidy and breeding system rather than the specific effects of the hidden states, we fit a simplified version with 16 parameters.
% FIXME Rosana, please update about the parameters
The reduction in parameter space was achieved by fixing the rates for transitions among hidden states to be equal with rate $\alpha$, and fixing the transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model).
There are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).
%B: I can't tell, but it seems to me that this 26->16 parameter reduction makes it possible to underestimate the effect of these hidden states.Perhaps this complication is covered in the Discussion section. 
%R- It is possible, but adding all the parameters can also result in an underestimation of the hidden states contribution because it is not enough sample to accurately estimate transitions. In my simulations for diversification-free models, accurately estimating 10 parameters can only be achieved with trees with more than 500 tips. Because of that experience I think 16 parameters is already pushing really the inference machine hard. This happens because the phylogeny is not an independent sample so your ESS is much much smaller and you might end up over-parameterizing.

@@

Parameters for each of the ten models were estimated using RevBayes \citep{hoehna_2016}.
Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $595/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models. % FIXME see referee comment about reporting convergence, not runtime
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

@@

We calculated the marginal likelihood for each of the ten models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
Each stepping stone step was found by calculating 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).
The calculation of each marginal likelihood ran for 24 hours on a high-performance computational cluster. % FIXME: remove?

Using the marginal likelihood values, we calculated thirteen different Bayes factors.
Six compared the models of ploidy against one other ($D/P$ and $D/P+A/B$, each with or without diploidization), one compared the breeding system models ($I/C$ and $I/C+A/B$), and six compared the models with both traits ($ID/CD/CP$ and $ID/CD/CP+A/B$, each with or without diploidization) (\cref{table:bayesfactors}).
%
Other comparisons between these models are not valid because the input data are different under the different state space codings (\cref{figure:stateclassifications}B).
In mathematical terms, the $D/P$, $I/C$, and $ID/CD/CP$ state spaces are not `lumpable' with respect to one another \citep{tarasov_2019}. % E: I'm honestly still confused about whether this is true.  From ID/CD/CP, we can lump ID/CD to get D/P, and we can lump CD/CP to get I/C.  FIXME
Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_0,\ M_1) = \ln[ P(\mathbf{X} | M_0) - P(\mathbf{X} | M_1)]$.
There is `positive' support for $M_0$ when this value is more than 2, `strong' support when it is more than 6, and `very strong' support when it is more than 10 \citep{kass1995}.

@@

\begin{table}
\begin{tabular}{lcccccc} \toprule
% \begin{tabular}{@{}lcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
\bottomrule
\end{tabular}
\caption{
    The ten models and their marginal likelihoods.
    Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
    Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
\label{table:marginallike}
\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|c|c|c|}
\toprule
Ploidy Models & Breeding System Models & Ploidy and Breeding System Models \\ \midrule
{\begin{tabular}{lcccc}
            & 1 & 2 & 3 & 4 \\
1. D/P$+\delta$ & $\cdot$     &10.72 &    -37.24    &-31.94\\
2. D/P  &$\cdot$&$\cdot$ &    -47.97    &-42.66\\
\textbf{3. D/P+A/B$+\delta$}  &$\cdot$  & $\cdot$&    $\cdot$    & 5.30 \\
4. D/P+A/B  &$\cdot$& $\cdot$ & $\cdot$&$\cdot$ \\
\end{tabular}
}  & 
{\begin{tabular}{lcc}
 & 5 & 6\\
5. I/C &$\cdot$ & -39.43 \\
\textbf{6. I/C+A/B} &$\cdot$& $\cdot$ \\
& & \\
& & \\
\end{tabular}
} & 
{\begin{tabular}{lcccc}
& 7 & 8 & 9 & 10\\
7. ID/CD/CP$+\delta$ & $\cdot$    &1.36 & -44.15&-40.95 \\
8. ID/CD/CP & $\cdot$ & $\cdot$ & -45.51 &-42.31 \\
\textbf{9. ID/CD/CP+A/B+$\delta$}& $\cdot$ & $\cdot$ &$\cdot$    & 3.2 \\ 
10. ID/P/CD+A/B & $\cdot$& $\cdot$&$\cdot$ & $\cdot$\\
\end{tabular}
}\\
\bottomrule
\end{tabular}
% E: We will probably have to change the layout here because it's not technically a table.  But good enough for now.
\caption{
    Bayes factors for model comparisons.
    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}B).
    Models are numbered in as \cref{table:marginallike}. % FIXME: they aren't, but it's a pain to change; change in prev table instead?
    Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
    Conventional thresholds for `strong' and `very strong' support are 6 and 10, respectively.
    The best model in each set is written in bold.
    In each case, it is the most complex model of the set.
}
\label{table:bayesfactors}
\end{table}

>>>>

Chromosome number data were obtained for all Solanaceae taxa in the Chromosome Counts Database \citep[CCDB;][]{rice_2015}, and the ca.~14,000 records were curated semi-automatically using the CCDBcurator R package \citep{rivero_2019}.
CCDB contains records from original sources that have multiple complex symbol patterns denoting multivalence, or irregularites of chromosome counts.
After a first round of automatic cleaning, we examined results by hand and corrected records as necessary.
Our hand-curated records were also contrasted against the ploidy dataset from \citet{robertson_2011}, original references therein, and against ploidy data in the C-value DNA dataset from \citet{bennett_2005}.
By comparing three different sources of information, we were able to code taxa as diploid, $D$, or polyploid, $P$.
For the majority of species, ploidy was assigned according to information from the original publications included in the  C-value DNA dataset \citep{bennett_2005}.
For taxa without ploidy information but with information about chromosome number, we assigned ploidy based on the multiplicity of chromosomes within the genus/family, or based on SI/SC classification.
For example, \textit{Solanum betaceum} did not have information about ploidy level but it has 24 chromosomes, so since $x=12$ is the base chromosome number of the genus \textit{Solanum} \citep{olmstead_2007}, we assigned \textit{S.~betaceum} as diploid. 
Additionally, because of the absence of SI polyploids (explained above and below), species known to be SI could be scored as diploid.
Species with more than one ploidy level were assigned the most frequent ploidy level recorded or the smallest ploidy in case of frequency ties.

@@

Matching our character state data to the largest time-calibrated phylogeny of Solanaceae \citep{sarkinen_2013} yielded 651 species with ploidy and/or breeding system information on the tree.
% FIXME: Rosana will update numbers around here TODO
Of these, 290 had information for both states.
The number of species in each combination of states is summarized in \cref{figure:stateclassifications}A.
We retained all 651 species in each of the analyses below because pruning away tips lacking breeding system in the ploidy-only analyses (and vice versa) would discard data that could inform the diversification models.
A total of 405 taxa without any information about breeding system or ploidy were excluded.
%Tips without trait data might not improve point estimation of diversification parameters while increasing uncertainty in the estimates of trait linked diversification
% R-This is very important but I have removed it because I think it is a full on paper by its own
%Including this many more species would have prohibitively slowed our analyses, especially those implementing the most complex models.

@@

% E: Rosana, please check this paragraph.  Did I get it correct?- R: Yes that is correct
To test if diversification rate heterogeneity is due only to a hidden trait and not the ploidy trait, we used character-independent diversification models \citep[CID,][]{beaulieu_2016}.
Parameters in this case are speciation and extinction rates for the hidden states ($\lambda_A,\ \lambda_B,\ \mu_A,\ \mu_B$), along with polyploidization ($\rho$) and transitions between hidden states ($(\alpha,\ \beta$).

@@

To parse out the effect of breeding system on diversification while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of the breeding system states into hidden states $A$ and $B$.
This $I/C+A/B$ breeding system and hidden state model has parameters for state-specific speciation ($\lambda_{I_A},\ \lambda_{I_B},\ \lambda_{C_A},\ \lambda_{C_B}$) and extinction ($\mu_{I_A},\ \mu_{I_B},\ \mu_{C_A},\ \mu_{C_B}$), for loss of SI ($q_{IC}$), and for transitions between hidden states ($\alpha$).
We also considered more complex models with different transition rates in each hidden state and asymmetric transitions between hidden states ($\alpha$ and $\beta$). % FIXME
Similar models were used by \citet{freyman_2017} used to study diversification in Onagraceae.

@@

Ploidy and breeding system might influence lineage diversification individually, but these two traits also have an intricate association as discussed in the Introduction.
Therefore, we considered a multi-state model that investigates the contribution of both traits and the allowable transitions between them.
This $ID/CD/CP$ ploidy and breeding system model has states for self-incompatible diploids $(ID)$, self-compatible diploids $(CD)$, and self-compatible polyploids $(CP)$; self-incompatible polyploids do not occur.
There are six parameters for state-specific diversification rates ($\lambda_{ID},\ \lambda_{CD},\ \lambda_{CP}$ for speciation; $\mu_{ID},\ \mu_{CD},\ \mu_{CP}$ for extinction) and three for transitions between states ($\rho_I,\ \rho_C$ for polyploidization transitions from $ID$ and $CD$ to $CP$, respectively; $q_{IC}$ for loss of self-incompatibility without polyploidization, from $ID$ to $CD$).
The total rate of loss of self-incompatibility, \ie transitions out of $ID$, is $q_{IC} + \rho_I$.
% E: Can we call it something other than q_IC?  It's a bit confusing to have the same parameter mean different things in different models.

% E: This paragraph is new.  Rosana, please check.
When including a new trait the question on whether 
the model with both breeding system and ploidy could provide a significantly better fit than the model with only ploidy was key.
Under all the statistical model comparison approaches, it is not valid to compare the $IC/CD/CP$ and $D/P$ model fits, as defined above, because they use different coding (i.e. (0,1) versus (0,1,2)). 
Similarly, the $IC/CD/CP$ and $I/C$ models cannot be compared statistically.
However, the evidence in the species and the tree are the same for both two-state and three-state models.
Instead, we define  and extend lumped models, as described first by \citet{tarasov_2019} for ontologies.  The lumped models proposed use the state space of the $IC/CD/CP$ model but constrain the rate parameters to mimic the behavior of the single-trait models. % should ref a figure

% E: We now have so many models to describe, I'm feeling like all this text may not be the best way.  I'm wondering if we can rely on figures/tables instead, and only write out the overall strategy. R- I feel the same but it might be very difficult to change at this point

Although the $ID/CD/CP$ model allows for two traits to potentially influence diversification, other unknown factors could as well.
We therefore added a hidden trait layer on top of our three-state model \citep[analogous to][]{caetano_2018, huang_2018}.
We fitted then $ID/CD/CP+A/B$ models, one with equal hidden rate $\alpha$ and equal transition rates between observed states to be independent of the hidden state (rates $\rho_I,\rho_C,\ q_{IC}$ as defined for the $ID/CD/CP$ model); second model were the hidden states are asymmetrical, and a third model were all parameters were different. For each of these models, there are additionally twelve diversification rate parameters ($\lambda_{ID_A},  \lambda_{ID_B},  \lambda_{CD_A}, \lambda_{CD_B}, \lambda_{CP_A}, \lambda_{CP_B}, \mu_{ID_A}, \mu_{ID_B}, \mu_{CD_A}, \mu_{CD_B},  \mu_{CP_A}, \mu_{CP_B}$).

@@

Parameters for each of the ten models were estimated using RevBayes \citep{hoehna_2016}.
Code for analyses and key results are available as Supplementary Information. % FIXME more permanent than \url{https://github.com/roszenil/solploidy}.
We accounted for incomplete sampling in all analyses by setting the probability of sampling a species at the present to $651/3000$ \citep[using the method of][]{fitzjohn_2009} since the Solanaceae family has approximately 3,000 species \citep{solsource}.
For all models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the  models. % FIXME see referee comment about reporting convergence, not runtime
Convergence and mixing of each MCMC chain was determined by ensuring the effective sample size of each parameter was over 200.

@@

We calculated the marginal likelihood for each of the models using fifty stepping stone steps under the methodology of \citet{xie_2010}, implemented in RevBayes \citep{hoehna_2016}.
Each stepping stone step was found by calculating at least 500 generations of burn-in followed by a total of 1,000 MCMC steps (\cref{table:marginallike}).

Using the marginal likelihood values, we calculated Bayes factors to argue five key biological and methodological questions:
\begin{enumerate}
\item Are hidden state necessary to add in this context?- Comparison of simple models against hidden state models
\item Is a second trait adding information to the diversification process?- Comparison of lumped models against IC/CD/CP models
\item Are diversification patterns only determined by hidden states and not the traits of interest?- Comparison of character independent models against hidden state 
\item Is there evidence for diploidization?- Comparison amongst models with and without diploidization
\item Do asymmetrical rates in the model affect the pattern of diversification? - Comparison amongst hidden states models with equal hidden rates and asymmetrical rates.
\end{enumerate}

Each model comparison is reported with a Bayes factor on the natural log scale: the comparison between models $M_0$ and $M_1$ is $BF(M_1,\ M_0) = \ln[ P(\mathbf{X} | M_1) - P(\mathbf{X} | M_0)]$.
There is `strong evidence'  for $M_1$ when this value is more than 2, `very strong' support when it is more than 6, and `extreme' support when it is more than 10 \citep{kass1995}.

@@

\begin{table}
\begin{tabular}{lcccccc} \toprule
% \begin{tabular}{@{}lcccccc@{}} \toprule
%\multicolumn{4}{r}{Models} \\ \cmidrule(r){3-5}
%Model                     & Ploidy & Diploidization & Breeding  & Hidden & Num        & Marginal       \\
%                          &        &                & System    & State  & Parameters & Log-Likelihood \\ \midrule
%1. D/P                    & Yes    & No             & No        & No     & 5          & -1193.66\\
%2. D/P +$\delta$          & Yes    & Yes            & No        & No     & 6          & -1182.93 \\
%3. D/P+A/B                & Yes    & No             & No        & Yes    & 10         & -1150.99\\
%4. D/P+A/B+$\delta$       & Yes    & Yes            & No        & Yes    & 11         & \textbf{-1145.69}\\
%5. I/C                    & No     & No             & Yes       & No     & 5          & -1194.80 \\
%6. I/C+A/B                & No     & No             & Yes       & Yes    & 10         & \textbf{-1155.37}\\
%7. ID/CD/CP               & Yes    & No             & Yes       & No     & 9          & -1345.87\\
%8. ID/CD/CP+$\delta$      & Yes    & Yes            & Yes       & No     & 10         & -1344.50\\
%9. ID/CD/CP+A/B           & Yes    & No             & Yes       & Yes    & 15         & -1303.55 \\ 
%10. ID/CD/CP+A/B+$\delta$ & Yes    & Yes            & Yes       & Yes    & 16         & \textbf{-1300.35} \\
%\bottomrule
%\end{tabular}
%\caption{
   % The ten models and their marginal likelihoods.
    %Values in bold are for the best models within each class that are comparable (see \cref{table:bayesfactors}).
    %Abbreviations are D: diploid, P: polyploid, I: self-incompatible, C: self-compatible, A: one state of hidden trait, B: other state of hidden trait, $\delta$: diploidization.}
%\label{table:marginallike}
%\end{table}

\begin{table}
\addtolength{\tabcolsep}{-3pt}
\begin{tabular}{|c|c|c|}
\bottomrule
\end{tabular}
% E: We will probably have to change the layout here because it's not technically a table.  But good enough for now.
\caption{
    Bayes factors for model comparisons.
    Each of the three boxes contains models that can be compared with one another, based on the character states they include (see \cref{figure:stateclassifications}B).
    Models are numbered in as \cref{table:marginallike}. % FIXME: they aren't, but it's a pain to change; change in prev table instead?
    Bayes factors are reported on the natural log scale, so numbers greater than $+2$ mean that the model in the row label has `positive' support relative to the model in the column label; numbers less than $-2$ mean that model in the column label is the preferred one.
    Conventional thresholds for `strong' and `very strong' support are 6 and 10, respectively.
    The best model in each set is written in bold.
    In each case, it is the most complex model of the set.
}
\label{table:bayesfactors}
\end{table}
